\documentclass[a4paper,10pt,landscape,twocolumn]{scrartcl}

%% Settings
\newcommand\problemset{7}
\newcommand\worksession{Tuesday, 16 October 2018}
\newif\ifcomments
\commentsfalse % hide comments
%\commentstrue % show comments

%% Packages
\usepackage[english]{exercises}
\usepackage{wasysym}
\usepackage{hyperref}
\hypersetup{colorlinks=true, urlcolor = blue, linkcolor = blue}
\usepackage{enumitem}
\usepackage{graphicx}

%% Macros
\input{macros}



\begin{document}

\problems

\begin{exercise}[RSA Signatures]
\begin{subex}
Consider the plain RSA signature scheme with modulus $N = 85 = 5 \cdot 17$.
\begin{enumerate}
\item  Compute $\phi(N)$.
\item  Say the public exponent is $e = 5$. Find the private exponent $d$. (Hint: look at
the sequence $\phi(N)+1, 2\cdot \phi(N)+1, \ldots$ until you find a multiple of $e$.)
\item  Compute the signature on the message $3$.\\\textbf{Hint: } Note that $[3^4 \mod 85] = [81 \mod 85] = [(-4) \mod 85]$.
\end{enumerate}
\end{subex}
\begin{subex}
Consider a \emph{padded} RSA signature scheme, where to sign a message $m \in \{0, 1\}^{80}$
the signer chooses random $r$ with $r \| m < N$ (where $\|$ denotes concatenation),
computes $\sigma = [(r \| m)^d \mod N ]$, and outputs signature $\sigma$.
\begin{enumerate}
\item  How would verification be done?
\item  Is this scheme secure? If yes, give a 1-2 sentence explanation; if not, show
an attack.
\end{enumerate}
\end{subex}
\end{exercise}


\begin{exercise}[Not a PRF]
Consider the keyed function $H : \{0, 1\}^n \times \{0, 1\}^n \rightarrow \{0, 1\}^{2n}$ defined as: $H_k(x) = G(k) \oplus G(x)$, where $G: \{0, 1\}^n \rightarrow \{0, 1\}^{2n}$ is a pseudorandom generator.
\begin{subex}
Describe and formally analyze an explicit attack showing that $H$ is not a PRF.
\end{subex}
\begin{subex}
Is there a successful attack making a single query that distinguishes $H_k$ (for random $k$) from a random function $f : \{0, 1\}^n \rightarrow \{0, 1\}^{2n}$? Why or why not?
\end{subex}
\end{exercise}



\begin{exercise}[A randomized variable-length MAC from a PRF]

Let $F$ be a pseudorandom function. Show that the following MAC is insecure for variable-length messages. $\mathsf{Gen}$ outputs a uniform $k\in\{0,1\}^n$. Let $\langle i\rangle$ denote an $n/2$-bit encoding of the integer $i$.

To authenticate a message $m=m_1\|\dots\| m_{\ell}$, where $m_i\in\{0,1\}^{n/2}$, choose a uniform $r\gets\{0,1\}^n$, compute $t:=F_k(r)\oplus F_k(\langle 1\rangle\Vert m_1)\oplus\cdots\oplus F_k(\langle \ell\rangle\Vert m_{\ell})$ and let the tag be $( r,t)$.

\end{exercise}


\begin{exercise}[Cryptographic Mechanisms]
For each of the following, identify the most appropriate cryptographic
mechanism(s) (from among private-key encryption, pseudorandom generators, pseudorandom functions, message authentication codes, hash functions, public-key encryption, or digital signatures) for addressing the problem. Points will be deducted if you
list extraneous mechanisms. \textbf{Explain your answer in 1-2 sentences.}
\begin{subex}
A company wants to distribute authenticated software updates to its customers.
\end{subex}
\begin{subex}
A user wants to ensure secrecy of the files stored on his hard drive.
\end{subex}
\begin{subex}
A customer wants to send his credit card number (confidentially) to a merchant
over the web to complete a purchase.
\end{subex}
\begin{subex}
A general wants to send a message to a lieutenant, and wants to ensure both
confidentiality and integrity.
\end{subex}
\begin{subex}
A client wants to store a short record of a large file he uploads to a server, so that
the client can verify that the file has not been altered when it downloads the file
later.
\end{subex}
\begin{subex}
A user needs 1,000,000 random bits in order to run a simulation, but obtaining
truly random bits is expensive.
\end{subex}
\end{exercise}





\end{document}
